---
title: "Template for Aesthetic Violin Plot"
author: "Michi Aneez"
date: "02/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### R Markdown

## This template is only applicable after all data cleaning and processing has taken place. 
## For a full example of how the lead up to the data visualisation looks like, refer to 'Lab 5 Data Analysis'. 

## First create a "master" df with all cleaned data in correct columns
    #this df should be labelled pinfo (participant inf)
## Then rename variables/columns (and groups/conditions)
    #this will sort the labelling of the data visualisation
    #ie., if there are multiple conditions, then renaming them here will remove the need to use labs() at the very last stage of the process
      #but it does make it annoying to type because the full names will require "" around them. just remember to spell it all correctly!
    #IRL Example:
```{r}
pinfo <- pinfo %>%
  mutate(randomiser.rknp = dplyr::recode(randomiser.rknp, "pitch_low" = "Low Anchor","pitch_high" = "High Anchor")) %>%
  rename(Condition = randomiser.rknp)
```

## To visualise the mean/median point on the graph, create a new df
    #make sure to have the same variable names in the left most column
    #IRL Example:
```{r}
desc.vis <- pinfo %>%
  group_by(Condition) %>% 
  summarise(Response = mean(Response))
 ```
#note that since this data is taken from the original pinfo df, it will automatically have the correct names for the responding columns
#to calculate the median (instead of the mean, like the above example), change the function to median()



## To visualise errorbars (i.e. to display the standard deviation/ standard error), create a new df
  #again, be careful to have the same variable/column names as the df with the mean/median values
  #IRL Example:
```{r}
errbar_lims <- pinfo %>%
  group_by(Condition) %>% 
  summarise(mean = mean(Response), sd = sd(Response), upper = mean+sd, lower = mean-sd)
 ```
#note that since this data is taken from the original pinfo df, it will automatically have the correct names for the responding columns
#to calculate the se (instead of the sd, like the above example), calculate the se by dividing the sd by the square root of the sample. 
  #note that the calculation of the upper and lower limits will remain the same (just with se instead of sd); only the calculation of the se value itself [se = sd(dv)/sqrt(nsample)] will be more complicated. 


## To visualise the aesthetic and informative violin plot, use the following code, but replace the variable/column names to ones that a relevant to your current df.
   #IRL Example:
```{r}
ggplot(pinfo, #identify the df you want to use
       aes(x = Condition, #clarify what goes in the x axis
           y = Response, #clarify what goes in the y axis
           colour = Condition)) + #gives two different colours to the conditions (n.b.: doesn't change  error bar colours)
  geom_jitter(alpha = .3, width = .09) + #adds spread of all data points, alpha indicates transparency 
  geom_violin(alpha = .2, width = .3) + # adds violin plot layer
  geom_point(data = desc.vis, #specifies that the data for the geom_point() layer will come from another df
             size = 2, width = .3) + #indicates the size and width of the median/mean point
  geom_errorbar(data = errbar_lims, #specifies that the data for the geom_errorbar() layer will come from another df
                aes(x = Condition, #does pretty much the same as the aes() for the ggplot(). because it's aesthetics.
                    ymin = lower, #specifies which column from the errorbar df will be lower bound
                    ymax = upper, #specifies which column from the errorbar df will be upper bound
                    colour = Condition), #makes the diff conditions' errorbars have diff colours, like the rest of the graph
                inherit.aes = FALSE, #prevents the layer from inheriting aesthetics from the parent plot
                width = 0.04) + #edits width of errorbars because they can be thicc sometimes
   labs(y = "Amount of Money Donated out of £100", #specifies y axis title
        title = "Hypothetical Participant Donations across the High and Low Anchor Conditions") + #gives graph a title
  guides(color = "none", fill = "none") + #this removes the ugly legend/key to the side indicating the groups
  theme_minimal() #gives the beautiful clean minimal theme babyyyyy
```

#congrats, you've now got one ✨ aesthetic ✨ graph. 

## Take care to keep ALL the column names the same when making new df's, otherwise the geom_point() and geom_errorbar() layers will not be added to the parent ggplot! 
## This is because the values on the new df's need to be plotted on the axis of the parent df and hence need to be recognised by the parent df and ggplot. 

---

## So a graph displaying only two conditions is good enough...
## But what if there are  multiple levels to the independent variable and there are multiple conditions too?
## Do not fear, since you can still make a sexy graph!
### Note: the real life example used below is not from "Lab 5"  


## To differentiate the different levels within the experimental conditions/groups, then you will need to use facet_wrap()
    #IRL Example (from Lab 4), including the creating of the additional df's to supplement the data viz:
```{r}
means <- dflong %>% 
  group_by(group, test) %>%
  summarise(score = mean(score))
#created a df with the mean values

allerrbar_lims <- dflong %>%
  group_by(group, test) %>%
  summarise(mean = mean(score), sd = sd(score), upper = mean+sd, lower = mean-sd)
#created a df with the error bar upper and lower bound/limit values

ggplot(dflong, aes(group, score, color = group)) + #no need to actually put "x = " or "y = " if listen in correct order
  facet_wrap(~test) + #facet wrap separates parts of the data by a specified group (i.e. "test" in this example)
  geom_jitter(alpha = .2) +
  geom_violin(aes(fill = group), #fills the space within the violin plot
              alpha = .2) + #but make sure to set the alpha (transparency) so that the geom_jitter() points are visible!
  geom_errorbar(data = allerrbar_lims, #specifies that the data for the geom_errorbar() layer will come from another df
                aes(x = group, ymin = lower, ymax = upper, colour = group), 
                inherit.aes = FALSE, #prevents the layer from inheriting aesthetics from the parent plot
                width = 0.04) + 
  geom_point(data = means, size = 2) +
  labs(x = "Test", y = "Mean Recall") +
  guides(color = "none", fill = "none") + 
  theme_bw() # theme_bw() works better with facet_wrap() than with theme_minimal(). but that may just be with this dataset so play around with themes anyway to find one that works best your current data viz!
```
#voila, you're done! ✨ aesthetic ✨ graph version 2.0 is ready to roll.
